---
title: "R program to extract tables from a PDF file"
author: "Himanshu Poddar"
date: "20 May 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Extract multiple tables from a PDF file

###Firstly Let us install some packages

install.packages("rJava")
library(rJava) # load and attach 'rJava' now
install.packages("devtools")
devtools::install_github("ropensci/tabulizer", args="--no-multiarch")

#Load the library tabulizer
```{r}
library(tabulizer)
```

###Now we are ready to extract tables from your PDF reports.
```{r}
PATH = 'C:/Users/Himanshu Poddar/Desktop/datathon/Himachal/bilaspur (h.p.) Class - 3 (Mathematics)  Report Card.pdf'
```

##extract all the tables present inside the pdf file
```{r}
lst <- extract_tables(PATH, encoding="UTF-8") 
```

##extract specific tables present inside the pdf file using coordinates

```{r}
lst <- extract_tables(PATH, encoding="UTF-8",pages = 1,area = list(c(234.019,38.991,313.638,555.396))) #co-ordinates separated with comma
```

##CODE to extract all the tables in a pdf file
```{r, eval = FALSE}
#Load library
library(tabulizer)
#Directory path
D_path = "C:/Users/Himanshu Poddar/Desktop/datathon/Himachal/"
files = list.files(D_path)
#looping through all the pdf files
for(file in files) 
{
  path = paste(D_path,file,sep = "")
  df = extract_tables(path, encoding="UTF-8") 
  #print(df)
}
```

##CODE to extract only specific table using its coordinates
```{r, eval = FALSE}
#Load library
library(tabulizer)
#Directory path
D_path = "C:/Users/Himanshu Poddar/Desktop/datathon/Himachal/"
files = list.files(D_path)
#looping through all the pdf files
for(file in files) 
{
  path = paste(D_path,file,sep = "")
  df = extract_tables(path,pages = 1,area = list(c(234.019,38.991,313.638,555.396)))  #where area is the coordinate of the table
  print(df)
}
```

##Conclusion
Now since all the tables are returned in a dataframe we can use R appropriate function to extract main content out of it.